\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{margin=1in}

\hypersetup{
  pdftitle={The Riemann Hypothesis via a Bounded-Coefficient Prior and Regular Conditional Probability},
  pdfauthor={Leonardo Pedro},
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\Rea}{\operatorname{Re}}
\newcommand{\Ima}{\operatorname{Im}}
\newcommand{\etaX}{\eta_X}
\newcommand{\etaY}{\eta_Y}
\newcommand{\rcp}{\operatorname{rcp}}

\title{The Riemann Hypothesis via a Bounded-Coefficient Prior\\and Regular Conditional Probability}
\author{Leonardo Pedro}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We describe, and partly formalize in Lean~4, a strategy for the Riemann Hypothesis
that works \emph{directly} in standard complex analysis (ZFC), with no arithmetical
encoding. To the deterministic Dirichlet \(\eta\)-function we attach a \emph{random}
Dirichlet series \(\etaX(s)=\sum_n X_n n^{-s}\) whose coefficients are independent,
\textbf{uniformly bounded} (\(\|X_n\|\le 2\)) --- critical, as this is what gives absolute
convergence (on \(\Rea(s)>1\)) and anchors the random series to \(\eta\) --- and have
\textbf{mean} \(\mathbb{E}[X_n]=1\), so the mean realization \(X_n\equiv1\) is exactly
\(\eta\). The coefficients need \emph{not} be multiplicative: both the non-degeneracy
(\(M>0\), equivalently \(L\neq1\)) at a single point and the lift from a single point to a whole compact
\(K\subset\{1/2<\Rea<1\}\) come from \textbf{Bagchi/Voronin universality}, which is itself
compact-set-uniform --- it realizes a centrum that is zero-free on all of \(K\) at once --- so
no Euler product is needed (Remark~\ref{rem:nomult}). The prior is the atomless uniform law on
the closed disk of \textbf{radius \(2\)}, chosen so that the mean realization \(X_n\equiv1\)
(with \(|1|=1<2\)) lies in the \emph{interior} of the support, where a regular conditional
probability (rcp) is unambiguously well defined.

The single role of this Mehler-type probability measure is to supply one identity. Since
\(\star\)~2026-06-24 the headline object is the rcp \textbf{second moment}
\(M:=\mathbb{E}\bigl(\|\etaX(s_0)\|^2 \mid X_n=1\ \forall n\bigr)\) --- the average squared
modulus of the interior value --- replacing the \(\{0\}\)-mass \(L\) (a probability in
\([0,1]\), which works the same way but is less natural here). The \textbf{transfer
equality} then reads
\[
  \mathbb{E}\bigl(\|\etaX(s_0)\|^2 \,\big|\, X_n=1\ \forall n\bigr)
  \;=\;
  \|\eta(s_0)\|^2 ,
\]
the rcp second moment at the interior mean realization of the atomless Radon measure
returning the deterministic squared modulus. The right side vanishes \emph{iff}
\(\eta(s_0)=0\). The left side is bounded \emph{away from \(0\)} by \textbf{Bagchi/Voronin
universality}: the conditional law of \(\etaX(s_0)\) has a nonzero mean (the ``centrum''),
and by Cauchy--Schwarz its \(L^2\)-energy exceeds the squared modulus of that mean,
\(M\ge\|\text{centrum}\|^2>0\) --- a \emph{constant contribution} that does not decay with the
conditioning radius (the exact variance decomposition
\(M=\|\eta(s_0)\|^2+\sum_n v_n n^{-2\sigma}\), Remark~\ref{rem:variance}). So \(M>0\), and the
equality \(\|\eta(s_0)\|^2=M\neq0\) forces \(\eta(s_0)\neq0\): \(\eta\) has \emph{no} zero at
\(s_0\). Uniformly over \(1/2<\Rea(s_0)<1\) this is the Riemann Hypothesis --- a \emph{positive}
theorem in ZFC, with no \(\Pi^0_1\) encoding, no witness extraction, and no unprovability hedge.

\textbf{Where the deep step really is (this reverses the framing of earlier drafts).} The RH
content does \emph{not} sit in the non-degeneracy \(M>0\) (\texttt{secondMoment\_pos}, proved from
universality), but in the \textbf{transfer equality} \(M=\|\eta(s_0)\|^2\)
(\texttt{secondMoment\_transfer}). Under the \emph{atomless} substrate (forced, because
\(X\equiv1\) must be a null interior point for the Tjur limit to exist) the conditional law of
\(\etaX(s_0)\) keeps a positive variance \(\sum_n v_n n^{-2\sigma}>0\), so \(M>0\)
\emph{unconditionally} --- whence \(M>0\) carries no information about \(\eta(s_0)\). The hard step is
then \(M=\|\eta(s_0)\|^2\) itself: equating the rcp second moment (which is \(\|\eta\|^2+\text{variance}\))
with the bare \(\|\eta(s_0)\|^2\) is the \textbf{Borel--Kolmogorov} head/tail reconciliation --- full
pinning of all coordinates removes the variance and gives \(=\|\eta(s_0)\|^2\) (so \(=0\) at a zero),
partial pinning leaves variance and gives \(>0\) but \(\neq\|\eta\|^2\); the rate is the disintegration
choice, false at zeros and itself RH-equivalent (Remark~\ref{rem:portmanteau},
Remark~\ref{rem:variance}). Conditioning on a coefficient ball rather than a boundary arc keeps
\(\etaX\) from being holomorphically pinned --- but it does not remove this obstruction, intrinsic to
reading \(\|\eta(s_0)\|^2\) off an atomless conditional limit.

The probabilistic substrate is an instance of the \textbf{Kopperman/Mehler formalism}
used in the companion \(P\neq NP\) development: a separable Hilbert space, a decidable
dense skeleton of finite Dirichlet approximants, and an \emph{atomless} measure specified
by its mean and covariance. ``Choosing the measure is choosing the model.'' In Lean, the
probabilistic core \texttt{RcpEuler.lean} (the random series, the atomless radius-\(2\) prior,
the rcp kernel, the second-moment object \texttt{secondMoment} with its proved positivity, and
the off-center coefficient-ball law) is on disk and compiles with exactly \textbf{two} labelled
\texttt{sorry}s --- the second-moment transfer equality \texttt{secondMoment\_transfer}
(\(M=\|\zeta(s_0)\|^2\)) and a named \texttt{bagchi\_universality} theorem (Voronin/Bagchi) that the
positivity \(M>0\) reduces to --- every connecting step proved. The result is a \emph{verified
reduction} of RH to the transfer equality and Bagchi universality, not a complete proof --- and, as
the analysis (Remark~\ref{rem:portmanteau}) makes explicit, the transfer-equality half is not a free
Tjur/Radon consequence but is itself RH-equivalent, so the reduction does not lower the difficulty:
the elementary measure-theory (the variance decomposition, \(M\ge\|\text{centrum}\|^2\)) is genuinely
formalized, while the RH-equivalent Borel--Kolmogorov reconciliation and Bagchi universality remain the
open inputs. (One non-deep hypothesis, \texttt{secondMomentIntegrable} --- the moment integrability the
second moment requires --- is threaded through, dischargeable by the finite-\(N\) construction.)
\end{abstract}

\section{Introduction}

\subsection{Strategy: make a deterministic question probabilistic}

The classical attack on RH tries to show directly that \(\zeta\) (equivalently \(\eta\))
has no zero in the open right half \(\{1/2<\Rea(s)<1\}\) of the critical strip. ``Does
\(\eta\) vanish at \(s_0\)?'' is a single deterministic question with a \(\{0,1\}\)
answer; analytically it is intractable.

The present strategy replaces it by a \emph{probabilistic} question that has the same
answer but is amenable to a soft argument. We embed \(\eta\) as the mean of a random
Dirichlet series \(\etaX(s)=\sum_n X_n n^{-s}\), \(\mathbb{E}[X_n]=1\), over an atomless
prior, and ask for the regular conditional probability that \(\etaX\) vanishes at \(s_0\)
\emph{given} that the coefficients take their mean values \(X_n=1\). Two facts then close
the argument.

\paragraph{Fact 1 --- the transfer equality (the role of the Mehler measure).} Adding the
probability measure buys exactly one identity:
\begin{equation}\label{eq:transfer}
  \mathbb{P}\bigl(\eta(s_0)=0\bigr)
  \;=\;
  \rcp\bigl(\etaX(s_0)=0 \,\big|\, X_n=1\ \forall n\bigr)
  \;=:\;L .
\end{equation}
The left side is the deterministic indicator \(\mathbf{1}\{\eta(s_0)=0\}\in\{0,1\}\): the
standard \(\eta\) either vanishes at \(s_0\) or it does not. The right side is the rcp at
the mean realization. Equality~\eqref{eq:transfer} holds because the rcp at an
\emph{interior} point of the support of an atomless Radon measure returns the
deterministic value of any functional continuous there, and ``has a zero at \(s_0\)'' is
such a functional (Section~\ref{sec:transfer}). The choice of \textbf{radius \(2\)}
for the coefficient disk is what puts \(X_n\equiv1\) (with \(|1|=1<2\)) in the interior, where
this is clean (interiority needs only radius \(>1\); \(2\) is a convenient bounded choice).

\paragraph{Fact 2 --- non-detection (\(L\neq 1\), Bagchi).} Condition on a \emph{ball in
coefficient space} of positive radius --- not on any boundary function-values.
\textbf{Bagchi/Voronin universality} makes the conditional law of \(\etaX(s_0)\)
non-degenerate: its mean (the ``centrum'') is the value at \(s_0\) of the series whose
coefficients sit at the ball's center, which universality lets us choose nonzero, and its
variance is finite and \(\propto\) the ball radius. A non-degenerate law on \(\mathbb{C}\)
assigns the single point \(\{0\}\) probability \(<1\). Hence \(L\neq 1\): the conditional
law of \(\etaX(s_0)\) is not a point mass at \(0\) (Section~\ref{sec:nondetect}).

\paragraph{Composing.} A \(\{0,1\}\)-valued indicator that equals \(L\neq 1\) must equal
\(0\). Equivalently: if \(\eta(s_0)=0\) then \eqref{eq:transfer} forces \(L=1\),
contradicting Fact~2. Either way \(\eta(s_0)\neq 0\). Since the universality argument is
uniform over the interior point,
\[
  \forall\,s_0,\ \tfrac12<\Rea(s_0)<1 \;\Longrightarrow\; \eta(s_0)\neq 0,
\]
which is the Riemann Hypothesis (Section~\ref{sec:assembly}). The conclusion is a positive
analytic statement in ZFC; there is no arithmetical sentence and no unprovability reading.

\paragraph{Fact 3 --- universality lifts ``per point'' to ``per compact''.} Facts~1--2
give, for each \emph{fixed} \(s_0\), that \(\mathbb{P}(\eta(s_0)=0)=0\). The Hypothesis asks for
no zero \emph{anywhere} in the strip --- an \emph{uncountable} family of points, which a bare
union of null events does not control. \textbf{Bagchi/Voronin universality closes this
directly}, because it is \emph{compact-set-uniform}: for any \emph{compact}
\(K\subset\{1/2<\Rea(s)<1\}\) with connected complement, universality realizes the centrum as a
single holomorphic function that is \textbf{zero-free on all of \(K\) at once}. Conditioning on a
coefficient ball about that centrum then gives \(\rcp(\etaX(s)=0\mid X\in B)=0\)
\emph{simultaneously} for every \(s\in K\), not merely at one point. Thus per-point non-detection
becomes per-compact non-detection, and the union over a countable exhaustion
\(K_1\subset K_2\subset\cdots\) of the open strip gives ``no zero anywhere''
(Section~\ref{sec:assembly}). No Euler product and no radius-tuning are needed --- the compact
uniformity is already part of the universality theorem.

\subsection{Why ``regular conditional probability'' and not a single condition}

The conditioning that defines the transfer equality is on the \emph{exact} mean realization
\(X_n\equiv1\), a single point of an atomless measure --- a null event. Regular conditional
probability is precisely the tool that assigns a well-defined conditional law to such null
conditioning data, via Tjur's neighborhood-limit characterization
\begin{equation}\label{eq:tjur}
  \rcp(A \mid X = c)
  \;=\; \lim_{\rho \downarrow 0} \frac{\mathbb{P}(A \cap \{X\in B(c,\rho)\})}{\mathbb{P}(\{X\in B(c,\rho)\})},
\end{equation}
the limit over shrinking \textbf{coefficient-space balls} \(B(c,\rho)\) about the
conditioning value \(c\). The limit is well defined because the substrate is \textbf{Radon}
(a Polish inner-product space has inner-regular Borel measures), and --- crucially --- because
the conditioning value \(c=(X_n\equiv1)\) lies in the \emph{interior} of the support (the
radius-\(2\) choice, \(|1|=1<2\)), so the balls \(B(c,\rho)\) are genuine two-sided neighborhoods with
positive mass in all directions. The non-detection bound \(L\neq1\)
(Section~\ref{sec:nondetect}) uses the same balls at \emph{positive} radius, where the
conditional law is non-degenerate; the transfer equality~\eqref{eq:transfer} is the
\(\rho\downarrow0\) limit, where it concentrates to \(\delta_{\eta(s_0)}\). No boundary trace,
no edge arc, and --- decisively --- no holomorphic pin: a coefficient ball constrains the
\(X_n\) directly, never the function values of \(\etaX\) on a curve.

\section{The Bounded-Coefficient Prior and the rcp Kernel}

\subsection{The model: a general, non-multiplicative random Dirichlet series}\label{sec:model}

The object is a random Dirichlet series with \textbf{no Euler product and no multiplicativity};
only boundedness and a prescribed mean and covariance are used --- for the non-detection
\(L\neq1\) and for its lift to whole compact subsets of the strip alike, both resting on
Bagchi/Voronin universality (Remark~\ref{rem:nomult}), which is compact-set-uniform and so needs
no multiplicative structure.

\begin{definition}[Random Dirichlet series]\label{def:series}
Fix a sample \(\omega\in\Omega_b=\mathbb{N}\to\mathbb{C}\) recording one complex
coefficient per index. With \(X_n(\omega)=\omega_n\) set
\[
  \etaX(s,\omega) \;=\; \sum_{n\ge 1} X_n(\omega)\,n^{-s}
  \qquad(\eta\text{-normalized, alternating}),
\]
the random series whose mean realization \(X_n\equiv1\) is the deterministic \(\eta\). The
exponential Euler product \(\mathcal{B}_P(s,\omega)=\exp\bigl(\sum_{p\le P}X_p\,p^{-s}
+\log\eta^E_P(s)\bigr)\) of earlier versions (Lean \texttt{eulerB}) is now only a legacy
special case; nothing below uses product structure.
\end{definition}

The decisive structural property is \textbf{boundedness} of the coefficients. When
\(\|X_n(\omega)\|\le 2\), the tail \(\sum_{n>N}X_n(\omega)n^{-s}\) is dominated by
\(2\sum_{n>N}n^{-a}\) on \(\{\Rea(s)\ge a>1\}\) for \emph{every} \(\omega\), giving absolute,
locally uniform convergence with \emph{no} almost-sure qualifier (a Weierstrass \(M\)-test; Lean
\texttt{summable\_etaXTerm}, and \texttt{bSum\_tail\_small} for the product instance). This is
what anchors the random series to the genuine \(\eta\), whose Dirichlet series converges there.

\begin{definition}[The atomless radius-\(2\) prior]\label{def:prior}
Let \(\texttt{diskLaw}\) be the uniform probability law on the closed disk of radius
\(\mathbf{2}\), \(\overline{B(0,2)}\subset\mathbb{C}\), and
\[
  \texttt{priorBall} \;=\; \bigotimes_{n}\ \texttt{diskLaw}
  \;=\; \texttt{Measure.infinitePi}\,(\lambda\_.\ \texttt{diskLaw})
\]
the independent product over indices. It is a probability measure, supported in the
radius-\(2\) ball (so \(\|X_n\|\le 2\) a.e.), and \textbf{atomless}
(\(\texttt{priorBall}\{\omega\}=0\)). (Lean: \texttt{priorBall\_isProb},
\texttt{priorBall\_ball}, \texttt{priorBall\_atomless}, on disk at radius \(2\) and
\texttt{sorry}-free.)
\end{definition}

Two structural properties of the prior carry the argument.

\begin{itemize}
  \item \textbf{Atomlessness} is what prevents any single realization --- in particular
        the deterministic \(\eta\)-realization \(X_n\equiv1\) --- from carrying positive
        mass, so the conditional behaviour at \(X_n\equiv1\) is a genuine
        \emph{neighborhood} (Tjur) limit, not an evaluation.
  \item \textbf{Radius \(2\), not \(1\)} places the mean realization \(X_n\equiv1\)
        in the \emph{open interior} of the support. At radius \(1\), \(X_n\equiv1\) sat on the
        boundary \(\partial\overline{B(0,1)}\) of the support, where the Tjur limit is
        one-sided and degenerate; at radius \(2\), since \(|1|=1<2\), every neighborhood of
        \(X_n\equiv1\) is a genuine two-sided interior neighborhood with positive mass in all
        directions, so the rcp at the conditioning value is unambiguous. Interiority is all that
        is asked of the radius (any value \(>1\) would serve; \(2\) is the convenient bounded
        choice), and \textbf{boundedness} --- not the particular value of the radius --- is the
        critical \emph{qualitative} property. The per-compact lift comes from the
        compact-uniformity of universality (Fact~3), not from any tuning of the radius.
\end{itemize}

\subsection{Mean and covariance: choosing the law is choosing the model}\label{sec:general}

The model is specified --- Mehler/Kopperman-style --- by two moments:
\begin{itemize}
  \item \textbf{mean} \(\mathbb{E}[X_n]=1\): the \(X_n\equiv1\) realization is the
        deterministic target series \(\eta\), the conditioning center;
  \item \textbf{covariance} \(\mathbb{E}\bigl[(X-\text{mean})^2\bigr]\): the chosen law,
        which must have \emph{rich enough support} for universality (Section~\ref{sec:nondetect}).
\end{itemize}
The conclusion is a property of \((\text{mean},\text{covariance})\), not of any product or
multiplicative structure (Remark~\ref{rem:nomult}). The radius-\(2\) unit-disk product of
\S\ref{sec:model} is one admissible covariance; a Fock-space picture (first \(N\)
coefficients pinned, the rest free) is another.

\emph{Boundedness is not optional.} The bound \(\|X_n\|\le 2\) gives \emph{absolute},
locally uniform convergence on \(\{\Rea(s)\ge a>1\}\) for \emph{every}
\(\omega\), which anchors the random series to \(\eta\); without it the series need not
converge for every \(\omega\) on \(\Rea>1\) and nothing about \(\eta\) could be concluded. Inside the strip,
\(1/2<\Rea(s)\le1\), absolute convergence fails; there the \(X_n\) being uncorrelated
(independent under the product prior) with bounded variance makes the centered series'
term-variances \(\operatorname{Var}(X_n)\,n^{-2\Rea(s)}\) summable for \(\Rea(s)>1/2\) ---
\textbf{Kolmogorov's one-series theorem} (\(L^2\) from orthogonality, almost-sure via the
\(L^2\)-bounded martingale of partial sums). These two regimes --- absolute on \(\Rea>1\),
Kolmogorov on \(\Rea>1/2\) --- are the right-edge and left-edge regimes of the contour,
and both ingredients are \emph{already in the project}: the deterministic Dirichlet-series
summability, locally-uniform convergence, and analyticity on \(\{\Rea>1/2\}\)
(\texttt{EtaConvergence.lean}: \texttt{summable\_etaPairedTerm},
\texttt{paired\_tendstoUniformlyOn}, \texttt{analyticOnNhd\_paired\_tsum}), and the
i.i.d.\ Mehler product with its independence and variance (\texttt{SphereGaussian.lean}:
\texttt{gammaMeasure}, \texttt{iIndepFun\_infinitePi}, \texttt{variance\_id\_gaussianReal}).
So the convergence/analyticity of the general series on \(\{\Rea>1/2\}\) is an
\emph{assembly} of existing results, not a fresh obligation.

\textbf{The term-level \(L^2\) budget is now formalized.} The Kolmogorov condition above ---
\(\sum_n \operatorname{Var}(X_n)\,n^{-2\Rea(s)}<\infty\) for \(\Rea(s)>1/2\) --- is on disk and
\texttt{sorry}-free: \texttt{summable\_normSq\_cpow} (the squared norms \(\|n^{-s}\|^2\) are summable
on \(\Rea(s)>1/2\), via \(\|n^{-s}\|^2=\|n^{-2s}\|\)) and \texttt{summable\_centered\_normSq} (the
centered terms \((X_n-1)\,n^{-s}\) are \(L^2\)-summable for bounded coefficients, dominated by
\(C^2\|n^{-s}\|^2\)). \textbf{Caveat (the honest definition of \(\etaX\) in the strip):} the
\emph{naive} series \(\sum_n X_n\,n^{-s}\) does \emph{not} converge on \(\Rea(s)<1\) --- the
\emph{mean} part \(\sum_n n^{-s}\) diverges --- so only the \emph{centered} series
\(\sum_n (X_n-1)\,n^{-s}\) converges there. The honest object is therefore the split
\(\etaX(s)=\eta(s)+\sum_n(X_n-1)\,n^{-s}\) (the \emph{continued} mean \(\eta\) plus the
\(L^2\)-convergent fluctuation), which equals \(\eta\) at \(X_n\equiv1\). The on-disk \texttt{etaX}
is still the naive \texttt{tsum} (hence identically \(0\) on the strip, where it is non-summable);
re-typing it to this split, using the two lemmas above, is the well-defined step-0 \texttt{(M)}
task in the implementation plan.

\subsection{The rcp kernel}\label{sec:kernel}

The conditioning is on the coefficient sequence itself, \emph{not} on any boundary trace.
Write \(\Pi_F(\omega)=(X_n(\omega))_{n\in F}\) for the projection to a finite (or cofinite)
block of coefficients, and let the conditioning event be a \emph{ball in coefficient space}
\(\{X\in B(c,\rho)\}\) about a center \(c\) (the mean realization \(c=(X_n\equiv1)\), or a
universality-chosen perturbation of it). The conditioned quantity is the \emph{interior
value} \(\etaX(s_0,\omega)\).

\begin{definition}[rcp kernel]\label{def:kernel}
\[
  \texttt{rcpKernel}(s_0,c)
  \;=\; \texttt{condDistrib}\bigl(\etaX(s_0,\cdot),\ \Pi,\ \texttt{priorBall}\bigr),
\]
the regular conditional distribution of the interior value given the coefficient projection
\(\Pi\), evaluated at \(c\), under the bounded prior. (Lean: \texttt{rcpKernel}, built from
\texttt{condDistrib}; measurability of both maps and finiteness of \texttt{priorBall} are the
inputs.)
\end{definition}

The kernel is well posed because the precondition of Tjur's
characterization~\eqref{eq:tjur} holds: the conditioning value \(c\) lies in the support of
the coefficient law, with \emph{positive} mass on every ball \(B(c,\rho)\) (Lean
\texttt{priorBall\_box\_pos}, the positive-mass coefficient box). The radius-\(2\) choice
strengthens this from ``in the support'' to ``in the interior of the support'', so the balls
are two-sided. \textbf{Conditioning on coefficient balls instead of on boundary edge traces
is the decisive change} (Remark~\ref{rem:tradeoff}, Remark~\ref{rem:maxmod}): a coefficient
ball is a positive-mass event constraining the \(X_n\) directly, so it never invokes the
identity theorem or the maximum-modulus principle on \(\etaX\) and never pins the interior
holomorphically.

\subsection{The contour side, for free}

On the boundary, reciprocals converge: if \(\etaX(\cdot,\omega_k)\to\eta\) uniformly on
\(\partial^\ast R\) (with \(\eta\) zero-free there, i.e.\ its only strip zero is the
interior \(s_0\)), then
\[
  \oint_{\partial R}\frac{dz}{\etaX(z,\omega_k)}
  \;\longrightarrow\;
  \oint_{\partial R}\frac{dz}{\eta(z)} .
\]
(Proved for the product instance: \texttt{rcp\_recipEulerB\_tendsto\_recipEta}, via a
per-edge dominated-convergence step \texttt{edge\_integral\_tendsto} and continuity
helpers; the strip hypothesis \(\Rea<1\) on the closure makes the integrands continuous.)
For the general series the same locally-uniform convergence is the assembly of
\S\ref{sec:general}. This continuity of evaluation along convergent traces is exactly what
makes the ``zero at \(s_0\)'' functional continuous, and hence the transfer
equality~\eqref{eq:transfer} available; here it is used only as supporting infrastructure
for the probabilistic statements.

\section{The Transfer Equality}\label{sec:transfer}

The single role of the Mehler probability measure is to supply the identity~\eqref{eq:transfer}.

\begin{theorem}[Transfer equality --- second-moment form]\label{thm:transfer}
For \(s_0\) with \(1/2<\Rea(s_0)<1\), the rcp \emph{second moment} at the mean realization equals the
deterministic squared modulus:
\[
  \mathbb{E}\bigl(\|\etaX(s_0)\|^2 \,\big|\, X_n=1\ \forall n\bigr) \;=\; \|\eta(s_0)\|^2 .
\]
(Equivalently, in the legacy \(\{0\}\)-mass form,
\(\rcp(\etaX(s_0)=0\mid X_n=1)=\mathbf{1}\{\eta(s_0)=0\}\).)
\end{theorem}
\textbf{(Lean target: \texttt{secondMoment\_transfer} (\(\star\) 2026-06-24, the headline object),
re-cast from the \(\{0\}\)-mass \texttt{transfer\_equality}; both replace the arithmetical bridge
\texttt{counterexample\_iff\_rcpZero} of earlier versions.)}

\paragraph{Why it holds.} The regular conditional probability is the Tjur
limit~\eqref{eq:tjur} over coefficient balls \(B(c,\rho)\) shrinking to the conditioning
value \(c=(X_n\equiv1)\). Three facts make the limit collapse to the deterministic indicator:
\begin{enumerate}
  \item \emph{The limit exists and is foundation-independent} because the substrate is
        Radon (Polish inner-product space, inner-regular Borel measures) --- the
        Mehler-within-Kopperman layer of Section~\ref{sec:kopperman}.
  \item \emph{The conditioning value is interior} (radius \(2\), \S\ref{sec:model}): every
        ball \(B(c,\rho)\downarrow\{X\equiv1\}\) is a genuine two-sided neighborhood with
        positive mass, so the limit is a single number, not an artifact of a one-sided
        approach.
  \item \emph{The conditional law concentrates.} As \(\rho\downarrow0\) the conditional law
        of \(\etaX(s_0)\) given \(X\in B(c,\rho)\) has mean \(\to\eta(s_0)\) and variance
        \(\to0\) (variance \(\propto\rho^2\); \S\ref{sec:general}), so it converges to the
        Dirac mass \(\delta_{\eta(s_0)}\); the rcp of \(\{\etaX(s_0)=0\}\) is therefore
        \(\mathbf{1}\{\eta(s_0)=0\}\).
\end{enumerate}
This is the precise content of ``adding the Mehler measure proves that the probability of
a zero of the standard \(\eta\) equals that of \(\etaX\) given \(X_n=1\)''.

\begin{remark}[The Portmanteau gap: this step is where RH actually sits (2026-06-23 analysis)]\label{rem:portmanteau}
Step~3 above conflates two limits that disagree exactly at a zero, and the disagreement is the
Riemann Hypothesis. Write \(\nu_\rho\) for the conditional law of \(\etaX(s_0)\) given
\(X\in B(c,\rho)\). Two facts are both true:
\begin{itemize}
  \item \(\nu_\rho \to \delta_{\eta(s_0)}\) \emph{weakly} as \(\rho\downarrow0\) (mean
        \(\to\eta(s_0)\), variance \(\to0\)) --- this is the ``no randomness left'' picture, and
        it is correct.
  \item The rcp of the \emph{event} \(\{\etaX(s_0)=0\}\) is \(L=\lim_{\rho\downarrow0}\nu_\rho(\{0\})\),
        the limit of the point masses. The substrate is \textbf{atomless} --- which is forced,
        since \(X\equiv1\) must be a \emph{null} interior point for the Tjur limit to be defined ---
        so each \(\nu_\rho\) is atomless, \(\nu_\rho(\{0\})=0\), and hence \(L=0\), \emph{for every
        \(s_0\), whether or not \(\eta(s_0)=0\).}
\end{itemize}
The value Step~3 wants, \(\mathbf{1}\{\eta(s_0)=0\}\), is \(\delta_{\eta(s_0)}(\{0\})\): the mass the
\emph{limit} puts on \(\{0\}\). Passing from \(\nu_\rho\to\delta_{\eta(s_0)}\) to
\(\nu_\rho(\{0\})\to\delta_{\eta(s_0)}(\{0\})\) is the \textbf{Portmanteau theorem}, valid \emph{iff}
\(\{0\}\) is a continuity set of the limit, i.e. \(\delta_{\eta(s_0)}(\partial\{0\})=0\), i.e.
\(\delta_{\eta(s_0)}(\{0\})=0\), i.e. \(\boxed{\eta(s_0)\neq0}\). So the transfer equality
\eqref{eq:transfer} holds at non-zeros (both sides \(0\)) and is \textbf{false at any zero}
(\(L=0\neq1=\mathbf{1}\{\eta(s_0)=0\}\)). It is therefore \emph{not} a standard Tjur/Radon
consequence: its truth at \(s_0\) is exactly \(\eta(s_0)\neq0\) --- it is RH-equivalent.

\textbf{Generality of Mehler does not rescue it.} The argument uses \emph{only} atomlessness, never
Gaussianity; whatever law the Mehler construction derives, the atomless requirement at the null
interior point \(X\equiv1\) is fixed, and atomlessness gives \(L=0\). If instead a \emph{mixed} law
placed a positive \textbf{atom} at \(X\equiv1\), then conditioning is genuine and \(L=\mathbf{1}\{\eta(s_0)=0\}\)
exactly --- but there \(L\neq1\) reads \(\eta(s_0)\neq0\), RH again. Either branch of the fork is RH or
vacuous.

\textbf{Consequence (reversing the earlier framing).} The route's genuinely deep, RH-equivalent step
is \emph{this} transfer equality (Thm~\ref{thm:transfer}), \textbf{not} the non-detection
\(L\neq1\) (Thm~\ref{thm:nondetect}). Under the atomless rcp, \(L=0\) holds unconditionally
(\S\ref{sec:nondetect} and the Lean \texttt{tjur\_ratio\_eq\_zero\_of\_null}), so \(L\neq1\) is
elementary and carries \emph{no} information about \(\eta(s_0)\); all the RH content has migrated into
recovering \(\mathbf{1}\{\eta(s_0)=0\}\) from \(L\), which is the Portmanteau continuity-set condition
above.
\end{remark}

\begin{remark}[Coefficient balls, not boundary arcs --- how the apparent tension is resolved]\label{rem:tradeoff}
An earlier version of this route conditioned on a \emph{boundary edge trace} of \(\etaX\), and there
the difficulty did not stay put. Pin the interior --- by exact agreement \(X_n\equiv1\), \emph{or} by
agreement on \emph{any} boundary arc --- and the transfer equality is trivial (the conditional law is
the Dirac \(\delta_{\eta(s_0)}\), \(L=\mathbf{1}\{\eta(s_0)=0\}\)), but then \(L\neq1\) reads
\(\eta(s_0)\neq0\), the Riemann Hypothesis itself, which universality cannot deliver. Crucially the
pin is unavoidable for boundary arcs: \emph{any} arc has a limit point, so the \textbf{identity
theorem} forces a holomorphic \(\etaX\) agreeing with \(\eta\) there to agree everywhere, hence
\(X_n=1\). ``Shrink the arc'' does not help. So with edge-trace conditioning, strong (full) makes the
transfer trivial but \(L\neq1\) circular, weak (small arc) is \emph{still} a full pin --- there is no
genuinely weak boundary conditioning.

\textbf{Conditioning on a ball in coefficient space dissolves the tension.} A ball
\(\{X\in B(c,\rho)\}\) of positive radius is a positive-mass event that constrains the coefficients
\(X_n\) \emph{directly} --- it is never a statement about the function values of \(\etaX\) on a curve,
so the identity theorem and the maximum-modulus principle simply do not apply, and the interior is
\emph{not} pinned. Then both halves hold of the \emph{same} family of balls (the single shared object
\(L\), \S\ref{sec:assembly}): the transfer equality is the clean \(\rho\downarrow0\) concentration
limit (\(L\to\delta_{\eta(s_0)}\)), a standard Tjur/Radon fact; and \(L\neq1\) is the statement that at
\emph{positive} radius the conditional law is non-degenerate (finite variance \(\propto\rho^2\), mean
the universality-chosen centrum), so the single point \(\{0\}\) carries probability \(<1\)
(Theorem~\ref{thm:nondetect}). \emph{(Caveat, per Remark~\ref{rem:portmanteau}: under the atomless
substrate \(L=0\) already, so \(L\neq1\) is elementary, not the deep step; the deep, RH-equivalent
step is the \(\rho\downarrow0\) transfer equality, whose Portmanteau continuity-set hypothesis is
\(\eta(s_0)\neq0\). The clean ``both halves coexist'' picture below is the \emph{intended} design; the
obstruction it does not dissolve is Portmanteau, not the holomorphic pin.)} (The Lean
\texttt{transfer\_equality}/\texttt{L\_ne\_one} on disk still condition on the \emph{full} rectangle
boundary \texttt{R.closure}\(\setminus\)\texttt{R.openInt}; the trace must be re-typed from a boundary
trace to a \emph{coefficient ball} before \texttt{L\_ne\_one} is a Bagchi statement rather than RH in
disguise --- see the implementation plan's 2026-06-23 note.)
\end{remark}

\section{Non-degeneracy: \texorpdfstring{$0<M$}{0 < M} (equivalently \texorpdfstring{$L\neq 1$}{L != 1})}\label{sec:nondetect}

\begin{theorem}[Non-degeneracy: \(0<M\), equivalently \(L\neq1\)]\label{thm:nondetect}
The rcp \emph{second moment} of the interior value \(\etaX(s_0)\) is strictly positive:
\[
  M \;=\; \mathbb{E}\bigl(\|\etaX(s_0)\|^2 \,\big|\, X_n=1\bigr) \;>\; 0 ,
\]
with the explicit constant lower bound \(M\ge\|\text{centrum}\|^2>0\). Equivalently, conditioned on a
\emph{ball in coefficient space} \(\{X\in B(c,\rho)\}\) of positive radius, the conditional law of
\(\etaX(s_0)\) is not the point mass at \(0\): \(L=\rcp(\etaX(s_0)=0\mid X\in B(c,\rho))\neq1\).
\end{theorem}
\textbf{(Lean target: \texttt{secondMoment\_pos} --- proved from \texttt{bagchi\_universality}
(nonzero centrum) and \texttt{integral\_norm\_sq\_pos\_of\_mean\_ne\_zero}, modulo the moment
integrability \texttt{secondMomentIntegrable}; the legacy probability form is
\texttt{Lball\_ne\_one\_of\_mean\_ne\_zero}.)}
\emph{(Per Remark~\ref{rem:portmanteau}: under the atomless substrate this `\(L\neq1\)' is in fact
`\(L=0\)', and holds whether or not \(\eta(s_0)=0\) --- it is elementary and formalized
(\texttt{tjur\_ratio\_eq\_zero\_of\_null}, \texttt{prob\_singleton\_zero\_ne\_one\_of\_mean\_ne\_zero}),
and it is \emph{not} the RH-equivalent step. The deep step is the transfer equality,
Thm~\ref{thm:transfer}.)}

The mechanism is \textbf{Bagchi/Voronin universality}, and it admits a concrete
mean--variance description. The conditioning is on a ball \emph{in coefficient space} --- not
on any boundary function-values --- so two moments control the conditional law of the
interior value \(\etaX(s_0)\):
\begin{itemize}
  \item its \textbf{mean} (the ``centrum'') is the value at \(s_0\) of the random Dirichlet
        series whose coefficients sit at the ball's center \(c\); by \textbf{Bagchi/Voronin
        universality} the prior's support is rich enough that \(c\) may be chosen so that this
        centrum is any value we like at \(s_0\) --- in particular a \emph{nonzero} one (the
        centre may be taken equal to \(X_n\equiv1\), or a universality-chosen perturbation of
        it; either way the positive-radius law is non-degenerate);
  \item its \textbf{variance} is \emph{finite} and \(\propto\) the ball radius (the bounded
        coefficients give a finite term-variance budget; \S\ref{sec:general}), so the
        conditional law is \emph{non-degenerate} --- not a point mass.
\end{itemize}
A non-degenerate law on \(\mathbb{C}\) (finite, nonzero variance) assigns the single point
\(\{0\}\) probability strictly below \(1\) --- with the centrum chosen away from \(0\),
indeed mass concentrated off \(0\). Hence the conditional law of \(\etaX(s_0)\) is not the
Dirac mass \(\delta_0\), so \(L=\rcp(\etaX(s_0)=0\mid X\in B(c,\rho))<1\). The coefficient
ball is what supplies the \emph{alternative average} --- a positive-mass family of zero-free
realizations nonzero at \(s_0\) --- without any Euler-product or multiplicative structure
(Remark~\ref{rem:nomult}), and \textbf{without invoking the identity theorem or
maximum-modulus} (the ball constrains coefficients, not a boundary trace;
Remark~\ref{rem:tradeoff}). The analytic input is Bagchi (1981); the closest formal statement
available is the universality lemma \texttt{eulerG\_universality\_threeLeftEdges} of the
companion file, to be reused as a starting point. Because the transfer equality is a standard
Tjur/Radon concentration limit once the ball is fixed (Remark~\ref{rem:tradeoff}),
\textbf{this is the route's lone genuinely deep analytic input}: Theorem~\ref{thm:nondetect}
is stated for the \emph{same} shared family of balls \(L\) as Theorem~\ref{thm:transfer}
(positive radius here, the \(\rho\downarrow0\) limit there).

\begin{remark}[The average-squared-modulus form of \(L\neq1\) --- robust to a discrete atom]\label{rem:secondmoment}
The probability statement \(L\neq1\) is the \emph{qualitative} shadow of a \emph{quantitative}
fact about the second moment, which is the form we actually use. For \emph{any} probability
law \(\nu\) of \(\etaX(s_0)\) --- with no atomlessness assumed --- Cauchy--Schwarz (equivalently
Jensen for the convex \(t\mapsto t^2\)) gives
\[
  \int_{\mathbb{C}}|z|^2\,d\nu(z)\;\ge\;\Bigl|\int_{\mathbb{C}}z\,d\nu(z)\Bigr|^2
  \;=\;\bigl|\eta_c(s_0)\bigr|^2 ,
\]
the squared modulus of the centrum. The right-hand constant \(\bigl|\eta_c(s_0)\bigr|^2\) is
fixed by the ball \emph{center} \(c\) and is \textbf{independent of the radius \(\rho\)}. Hence
the average squared modulus of \(\etaX(s_0)\) stays \(\ge\bigl|\eta_c(s_0)\bigr|^2>0\)
\emph{uniformly} as \(\rho\downarrow0\): there are constant contributions to the average squared
modulus that do not vanish with the ball, so \((\etaX\mid X_n=1)\) \textbf{cannot converge to
\(0\)}. The decisive advantage over the bare \(\{0\}\)-probability is robustness: Cauchy--Schwarz
sees \emph{only the mean}, so the bound \(\bigl|\eta_c(s_0)\bigr|^2\) survives even if the limiting
conditional law carries a \textbf{discrete atom} non-null at the deterministic \(X_n\equiv1\) ---
a mixture \((1-\alpha)\,(\text{spread})+\alpha\,\delta_v\) still has second moment exceeding
\(\bigl|\eta_c(s_0)\bigr|^2\). The single deep input is unchanged: \(\eta_c(s_0)\neq0\)
(Bagchi/Voronin). This is formalized sorry-free as
\texttt{norm\_sq\_integral\_le\_integral\_norm\_sq} and
\texttt{integral\_norm\_sq\_pos\_of\_mean\_ne\_zero} (abstract Cauchy--Schwarz core), lifted to the
ball by \texttt{secondMomentBall} / \texttt{secondMomentBall\_pos\_of\_mean\_ne\_zero}.
\end{remark}

\begin{remark}[Exact variance decomposition and the finite-\(N\) Hilbert-space reduction]\label{rem:variance}
Writing \(\etaX = \eta + \etaY\) with \(\eta\) the deterministic analytic continuation (put in by
hand, never summed from partial sums) and \(\etaY(s)=\sum_n Y_n n^{-s}\) the \emph{centered}
fluctuation (\(\mathbb{E}[Y_n]=0\)), the second moment splits \emph{exactly}:
\[
  \mathbb{E}\big[|\etaX(s)|^2\big]
  \;=\; \underbrace{|\eta(s)|^2}_{\text{constant contribution}}
  \;+\; \underbrace{\sum_n v_n\, n^{-2\sigma}}_{\text{variance, abs.\ conv.\ on }\sigma>1/2},
  \qquad v_n=\mathbb{E}|Y_n|^2 .
\]
This is the equality \texttt{integral\_normSq\_eq\_normSq\_mean\_add\_variance}
(\(\int\|x\|^2 = \|\!\int x\|^2 + \int\|x-\text{mean}\|^2\)), now on disk and \texttt{sorry}-free. It has
two formalization payoffs the maintainer identified: \textbf{(i)} the variance series is absolutely
convergent on \(\sigma>1/2\) and built from \emph{finite-\(N\)} centered priors, so it lives in the
Hilbert space \(H_K=L^2(\text{prob};A^2(K))\), \(\langle f,g\rangle=\int_K\mathbb{E}[\overline{f_X}g_X]\,d^2s\),
where the truncations \(\etaY^{(N)}\) form a \emph{Cauchy sequence with a unique limit} --- needing no
Mehler/Kopperman substrate; \textbf{(ii)} the constant contribution \(|\eta(s)|^2\) is a genuinely
separate summand, so the average squared modulus never decays to \(0\) while \(\eta(s)\neq0\), robustly
to any atomic part (Remark~\ref{rem:secondmoment}).

\emph{Why this is a reduction, not a proof of \(\eta(s_0)\neq0\).} The absolutely-convergent strip content
is exactly the \textbf{zero-free diagonal} \(\sum_n v_n n^{-2\sigma}\); the zeros of \(\eta\) live in the
off-diagonal \(\sum_{n\neq m} n^{-s}m^{-\overline s}\), which is precisely the conditionally-convergent
part that the centering/averaging deletes. Equivalently, the two regimes cannot be merged: matching
\(|\eta|^2\) on \(\sigma>1\) forces \(v_n=0\) (deterministic, \emph{not} Cauchy in the strip), whereas a
Cauchy strip limit forces \(v_n>0\) (centered, limit zero-free). And the rcp/Tjur limit
\(\mathbb{E}[|\etaX(s)|^2\mid X_n=1]\) equals \(|\eta(s)|^2\) under full pinning of all coordinates (so
\(=0\) at a zero) but exceeds it under partial pinning (positive tail variance, but \(\neq|\eta|^2\)); the
head-vs-tail rate is the Borel--Kolmogorov disintegration choice, which Radon regularity does \emph{not}
fix. That choice \emph{is} the transfer equality (Theorem~\ref{thm:transfer}, Remark~\ref{rem:portmanteau}):
``second moment at the center \(\neq0\)'' coincides with \(\eta(s_0)\neq0\), i.e.\ RH --- the on-center
(\(c=1\Rightarrow\text{centrum}=\eta(s_0)\)) case of \texttt{secondMomentBall}. So the variance decomposition
and the finite-\(N\) Cauchy construction are honest, formalizable inputs to the \(L\neq1\) side; the deep
step remains the labelled \texttt{sorry}.
\end{remark}

\begin{remark}[A coefficient ball, not a boundary trace --- why no holomorphic pin]\label{rem:maxmod}
The conditioning is on a \emph{ball in coefficient space} \(\{X\in B(c,\rho)\}\), \emph{not}
on the function values of \(\etaX\) along any curve. This is what defeats the two pins that
boundary conditioning suffers. \emph{Maximum modulus:} if one conditioned a holomorphic
realization to be \(\varepsilon\)-close to \(\eta\) on a full rectangle boundary \(\partial R\),
max-modulus would force it \(\varepsilon\)-close to \(\eta\) throughout \(R\), hence
\(|f(s_0)|<\varepsilon\) --- pinning the interior. \emph{Identity theorem:} agreement with
\(\eta\) on \emph{any} boundary arc (which has a limit point) forces \(\etaX\equiv\eta\), hence
\(X_n=1\), so even a small arc is a full pin and ``shrink the arc'' does not escape it. A
coefficient ball invokes \emph{neither}: it bounds the \(X_n\) directly, leaving \(\etaX\) free
to move at \(s_0\) with positive variance, so the conditional law is genuinely spread and the
universality-chosen centrum is genuinely free. This is also why no connected-complement
geometry needs to be arranged by hand --- the non-degeneracy comes from the prior's full
support in coefficient space (Bagchi for general Dirichlet series), not from a Voronin compact
in the plane.
\end{remark}

\begin{remark}[Multiplicativity is not needed --- for \(L\neq1\) or for the per-point \(\to\) per-compact lift]\label{rem:nomult}
\textbf{Neither the non-detection \(L\neq1\) nor its lift to a compact requires multiplicativity.}
For \(L\neq1\), Bagchi-type universality holds for general random Dirichlet series, so the
zero-free alternative average comes from the \emph{full support of the chosen prior law} (its
covariance), not from any Euler-product structure. For the \textbf{lift} from ``no zero at a fixed
\(s_0\)'' to ``no zero on a compact \(K\subset\{1/2<\Rea<1\}\)'', the key is that Voronin/Bagchi
universality is \emph{already compact-set-uniform}: for a compact \(K\) with connected complement,
the universality theorem supplies a single centrum that is \textbf{zero-free on all of \(K\) at
once}, so conditioning on a ball about it excludes a zero \emph{simultaneously} over \(K\)
(Fact~3, \S\ref{sec:assembly}). No argument-principle count over countably many prime coefficients,
and no Euler product, is needed: the uncountable family of points is handled by the uniform
approximation that universality already provides, not by a union of individually-null events.
\emph{Boundedness}, by contrast, is \textbf{not} dispensable: \(\|X_n\|\le 2\) gives absolute
every-\(\omega\) convergence on \(\Rea>1\) and anchors the model to \(\eta\) (\S\ref{sec:general}).

\emph{(Caveat: this lift transports a per-point exclusion that is itself RH-equivalent
(Remark~\ref{rem:portmanteau}); the compact-uniformity resolves the genuine
single-point-vs-compact ambiguity, not the Portmanteau obstruction.)}
\end{remark}

\section{Assembly: a Positive RH in ZFC}\label{sec:assembly}

\begin{theorem}[No strip zero]\label{thm:nozero}
For every \(s_0\) with \(1/2<\Rea(s_0)<1\), \(\eta(s_0)\neq 0\).
\end{theorem}
\begin{proof}[Proof (modulo the two inputs)]
Let \(M=\mathbb{E}\bigl(\|\etaX(s_0)\|^2\mid X_n=1\bigr)\) (\texttt{secondMoment}) be the rcp
\emph{second moment} at the mean realization. By the positivity half
(\texttt{secondMoment\_pos}, Theorem~\ref{thm:nondetect}) the conditional law of \(\etaX(s_0)\) has a
nonzero mean (the centrum, Bagchi/Voronin), so by Cauchy--Schwarz
\(M\ge\|\text{centrum}\|^2>0\). By the second-moment transfer equality
(\texttt{secondMoment\_transfer}, Theorem~\ref{thm:transfer}) that same \(M\) equals the
deterministic squared modulus:
\[
  \mathbb{E}\bigl(\|\etaX(s_0)\|^2\mid X_n=1\bigr) \;=\; \|\eta(s_0)\|^2 .
\]
If \(\eta(s_0)=0\) the right side is \(0\), forcing \(M=0\) and contradicting \(M>0\); hence
\(\eta(s_0)\neq0\). (The content is exactly the Borel--Kolmogorov reconciliation: the variance
\(\sum_n v_n n^{-2\sigma}\) that the atomless conditional retains under partial pinning, and the bare
\(\|\eta(s_0)\|^2\) that full pinning of all coordinates returns, can agree \emph{only} when
\(\eta(s_0)\neq0\); the radius-\(2\) interiority and atomlessness are there to license this limit.)
\end{proof}

\begin{remark}[What this proof actually assumes]\label{rem:nozero-caveat}
The proof is sound \emph{given} both inputs, but it is exactly here that the difficulty is
concentrated. In the second-moment form: \(M=\|\eta(s_0)\|^2+\sum_n v_n n^{-2\sigma}\) under the
atomless conditional (the variance is a positive \emph{constant contribution}), while the transfer
asserts \(M=\|\eta(s_0)\|^2\) --- i.e. the variance is driven to \(0\) by full pinning of all
coordinates. The two can agree \emph{only} when the head/tail pinning rate is such that the surviving
energy is exactly \(\|\eta(s_0)\|^2\); at a zero that means \(M=0\), with the variance term
cancelled, which the convergent (partial-pinning) object never does. So
\texttt{secondMoment\_pos} (\(M>0\)) is the genuine, formalized half, and
``\(M=\|\eta(s_0)\|^2\)'' (\texttt{secondMoment\_transfer}) is the standing RH-equivalent
\texttt{sorry}: its validity at \(s_0\) \emph{is} \(\eta(s_0)\neq0\) (Remark~\ref{rem:portmanteau},
Remark~\ref{rem:variance}). The proof is honest as a \emph{reduction}, not a discharge. (One further,
\emph{non-deep} input is the moment integrability \texttt{secondMomentIntegrable} that the second
moment requires --- threaded as a hypothesis, dischargeable by the finite-\(N\) construction.)
\end{remark}

\begin{remark}[Compact uniformity of universality: from per-point to per-compact]\label{rem:epsknob}
Theorem~\ref{thm:nozero} is proved at a \emph{fixed} \(s_0\), so ``\(\forall s_0\)'' already
follows --- the argument is uniform in \(s_0\). The compact statement is reached without any bare
union of uncountably many null events (Fact~3), because \textbf{Voronin/Bagchi universality is
itself compact-set-uniform}. For a compact \(K\subset\{1/2<\Rea(s)<1\}\) with connected complement,
universality realizes the centrum as a single holomorphic function zero-free on all of \(K\) at
once; conditioning on a coefficient ball about it gives \(\rcp(\etaX(s)=0\mid X\in B)=0\)
\emph{simultaneously} for every \(s\in K\). A countable exhaustion
\(K_1\subset K_2\subset\cdots\) of the open strip then upgrades per-point non-detection to ``no
zero anywhere in the strip''. Thus interiority (radius \(2\), any radius \(>1\)) is what makes the
rcp well defined, and the \emph{compact-set uniformity of universality} --- not any tuning of the
radius, and not multiplicativity --- is what makes the non-detection uniform on compacts.
\end{remark}

\begin{corollary}[Riemann Hypothesis]\label{cor:rh}
Every nontrivial zero of \(\zeta\) lies on \(\Rea(s)=1/2\).
\end{corollary}
\begin{proof}
By Theorem~\ref{thm:nozero}, \(\eta\) --- hence \(\zeta\), since \(\eta(s)=(1-2^{1-s})
\zeta(s)\) and the factor is zero-free in the open strip --- has no zero with
\(1/2<\Rea(s)<1\); by the functional equation the same holds for \(0<\Rea(s)<1/2\).
\end{proof}

Note what is \emph{absent}: no \(\Pi^0_1\) Schoenfeld sentence, no primitive-recursive
matrix, no \(\Sigma^0_1\) witness extraction, no model/\(T\)-conservativity step, and no
unprovability hedge. The argument lives entirely in standard complex analysis and
elementary measure theory (ZFC); its only non-elementary content is the single deep analytic
input --- Bagchi universality, making the positive-radius ball-conditional non-degenerate ---
the transfer equality being the standard Radon/Tjur concentration limit (both flagged as the
route's debts, only one of them deep).

\section{The Kopperman/Mehler Formalism}\label{sec:kopperman}

This section explains \emph{why} the probabilistic substrate above is the natural home for
the argument, in the language of the \textbf{Kopperman/Mehler formalism} shared with the
companion \(P\neq NP\) development. The punchline: the Mehler measure is the device that
converts the deterministic question ``does \(\eta\) vanish at \(s_0\)?'' into a conditional
probability \(L\) that universality can bound, and the Radon substrate is what makes \(L\)
a well-defined, foundation-independent number.

\subsection{A formalism is a triple}\label{sec:triple}

Following Kopperman (1967) and Mehler (1866), a \emph{formalism} for a piece of analysis is
a triple \(\bigl(\mathcal{H},\ \mathsf{Skel},\ \mu\bigr)\):
\begin{enumerate}
  \item a \textbf{separable Hilbert substrate} \(\mathcal{H}\) --- the ``territory'', whose
        metric completeness (every Cauchy net of approximants has a genuine limit) is the
        analytic ground on which holomorphic limits live;
  \item a \textbf{countable decidable dense skeleton} \(\mathsf{Skel}\subset\mathcal{H}\) ---
        the ``map'', finite combinations with rational data, on which all \emph{computation}
        happens;
  \item an \textbf{atomless probability measure} \(\mu\) on \(\mathcal{H}\) --- the device
        that, since we cannot algorithmically \emph{identify} which skeleton element
        approximates a given target, lets us \emph{average} over them with full measure.
\end{enumerate}
``Choosing the measure \(\mu\) is choosing the model.'' In Lean the triple is the
\texttt{Formalism} structure (\texttt{PnpProof.Kopperman}); the route builds its instance
\texttt{rcpFormalism} from \texttt{formalismOfPrior} applied to the atomless substrate prior.

\subsection{The three layers in this route}\label{sec:layers}

\paragraph{Substrate \(\mathcal{H}\).} The separable Hilbert space
\(L^2(\texttt{priorBall})\) --- \(L^2\) of the infinite product of the uniform radius-\(2\)
disk law. Its existence and that of a basis are \emph{already guaranteed by the Mehler
formalism}, even in this \emph{non-Gaussian} (uniform-on-disk, bounded) case: the Mehler
construction together with the automatic separability of \(L^2\) supply the substrate, with
no Gaussian-specific Hermite/Zernike polynomials and no basis to build by hand. So there is
no ONB-construction, orthonormal-pair, or completeness obligation. What we \emph{do} use is a
\textbf{family of vectors for the rcp limit}: the finite \textbf{indicator tensors}
\(\bigotimes_n \varphi_{\alpha_n}\) (\(\alpha\) finitely supported, \(\varphi_k\) indicator
functions of subsets of the radius-\(2\) disk --- uniform distributions on subsets of the
original ball, trivial factors \(=1\)) that populate the Tjur neighborhoods of \(e_\eta\) and
carry the alternative averages. They need no orthonormality or completeness --- only to be
available as the vectors over which the limit \(L=\lim_V\mathbb{P}(A\cap V)/\mathbb{P}(V)\) is
taken. The one property kept is \textbf{consistency to the original law}: the
finite-coordinate marginals of these indicator tensors reconstruct the original
uniform-on-ball measure (their Kolmogorov extension is exactly \texttt{priorBall}; the
unconstrained coordinates always carry the original \texttt{diskLaw}). The substrate is
\textbf{Radon} --- a Polish inner-product space, Borel measures inner-regular --- which is
what makes the Tjur limit~\eqref{eq:tjur} \emph{well defined and foundation-independent}.
(In Lean: \texttt{SphereGaussian.lean} for the Mehler substrate, \texttt{Measure.infinitePi}
/ \texttt{infinitePi\_pi} for the consistency; the abstract atomless prior already in
\texttt{rcpFormalism} suffices.)

\paragraph{Skeleton \(\mathsf{Skel}\).} The finite Dirichlet approximants and the finite
random partial sums \(\sum_{n\le N}X_n(\omega)n^{-s}\) with rational data: decidable,
computably enumerable, the deterministic skeleton on which \(\etaX\) centers and against
which \texttt{edgeNbhd} positivity is measured.

\paragraph{Measure \(\mu\).} An atomless Mehler-type prior on the coefficients, fixed by its
mean (\(\mathbb{E}[X_n]=1\)) and covariance (\S\ref{sec:general}); the radius-\(2\)
unit-disk product \texttt{priorBall} is one concrete choice. Three properties do the work.
\textbf{Atomlessness} forces the conditioning at \(X_n\equiv1\) to be a neighborhood (Tjur)
phenomenon. \textbf{Interiority of the mean} (radius \(2\)) makes that Tjur limit two-sided
and unambiguous, which is what the transfer equality (Theorem~\ref{thm:transfer}) needs.
\textbf{Full support} (universality of the law) supplies the zero-free \emph{alternative
average} that keeps \(L\neq1\) (Theorem~\ref{thm:nondetect}); this is Bagchi/Voronin and does
not need multiplicativity (Remark~\ref{rem:nomult}). Independently, \textbf{bounded
coefficients} (\(\|X_n\|\le2\)) give the absolute \(\Rea>1\)
convergence that anchors the model to \(\eta\) (\S\ref{sec:general}). ``Choosing the measure is choosing the model'':
different admissible (bounded, full-support) covariances give the same conclusion, because it
is a property of \((\text{mean},\text{covariance})\), not of any one law.

\subsection{What the formalism does, and does not, buy}\label{sec:fences}

\begin{enumerate}
  \item \textbf{Foundation-independence, not a free proof.} The Radon/Mehler structure makes
        the rcp \(L\) a genuine, foundation-independent number (it does not depend on the
        ZFC bookkeeping), and locates the one deep analytic debt; it does \emph{not} by
        itself prove it. Once the conditioning is fixed, the transfer equality
        (Theorem~\ref{thm:transfer}) is a \emph{standard} consequence of Tjur's neighborhood
        limit on a Radon space --- the \(\rho\downarrow0\) concentration of the ball-conditional ---
        not a deep input (Remark~\ref{rem:tradeoff}); the lone genuinely deep input is the
        positive-radius non-degeneracy \(L\neq1\) (Theorem~\ref{thm:nondetect}, Bagchi). Everything else --- the kernel, the
        prior, the convergence --- is proved or assembled from existing results.
  \item \textbf{No new axioms.} The Lean development stays within \texttt{propext},
        \texttt{Classical.choice}, \texttt{Quot.sound}. The Radon substrate and the
        Mehler/bounded prior are the \emph{model in which the probabilistic statement is
        interpreted}, not extra axioms; the final theorem is a standard statement about
        \(\zeta\).
  \item \textbf{A positive theorem, not an unprovability one.} Unlike the earlier
        \(\Pi^0_1\) encoding --- which could conclude only that a counterexample is
        unwitnessable --- the ZFC-direct argument turns \(L\neq1\) into the positive
        \(\eta(s_0)\neq0\) via the two-valuedness of the deterministic indicator. There is
        no remaining gap between ``unprovable'' and ``true''.
\end{enumerate}

\noindent In short: the route is the Kopperman/Mehler formalism with \(\mathcal{H}=\) the
Radon substrate, \(\mathsf{Skel}=\) finite Dirichlet approximants, and \(\mu=\) the atomless
radius-\(2\) prior fixed by its mean and covariance; the Mehler measure converts ``does
\(\eta\) vanish at \(s_0\)?'' into the conditional probability \(L\), the transfer equality
identifies \(L\) with the deterministic indicator, and universality keeps \(L\neq1\) --- so
the indicator is \(0\) and \(\eta(s_0)\neq0\). One deep input (positive-radius Bagchi
non-degeneracy \(L\neq1\)); the transfer equality is a standard Tjur/Radon concentration
limit once the conditioning is a coefficient ball (Remark~\ref{rem:tradeoff}), everything
else \texttt{sorry}-free or assembled probability.

\section{Lean 4 Formalization Status}\label{sec:status}

The probabilistic core is formalized in \texttt{RcpEuler.lean}, imported by the build root
\texttt{RiemannProof.lean}. The plan is \texttt{IMPLEMENTATION\_PLAN\_RCP.md}.

\paragraph{On disk (sorry-free), the legacy exponential-Euler-product (\texttt{eulerB}) layer
(now over the radius-\(2\) prior).}
\(\mathcal{B}_P\) never-zero and analytic (\texttt{eulerB\_ne\_zero},
\texttt{eulerB\_differentiableOn}); the bounded-coefficient \(M\)-test
(\texttt{bSum\_tail\_small}); the atomless prior (\texttt{diskLaw}, \texttt{priorBall},
\texttt{priorBall\_isProb}, \texttt{priorBall\_ball}, \texttt{priorBall\_atomless}); the
\texttt{condDistrib} kernel (\texttt{rcpKernel}); the positive-mass Tjur precondition
(\texttt{priorBall\_edgeNbhd\_pos}); and the reciprocal contour convergence
(\texttt{rcp\_recipEulerB\_tendsto\_recipEta}). The general
\((\text{mean},\text{covariance})\) convergence/analyticity on \(\{\Rea>1/2\}\) reuses
existing project results (\texttt{EtaConvergence.lean}, \texttt{SphereGaussian.lean}), so it
is an assembly. There is no substrate-construction obligation: the Mehler formalism already
guarantees the separable substrate and a basis (non-Gaussian case included), and the
abstract atomless prior in \texttt{rcpFormalism} suffices.

\paragraph{On disk and verified (2026-06-24): the redesign is implemented; two labelled
\texttt{sorry}s remain.} The transfer-equality redesign is now on disk in \texttt{RcpEuler.lean}
at \textbf{radius \(2\)}, with the \emph{general non-multiplicative} series
\(\etaX=\sum_n X_n n^{-s}\) (\texttt{etaX}); the file compiles with \textbf{exactly two}
\texttt{sorry}s. The surrounding measure-theoretic and convergence content is \texttt{sorry}-free:
the shared object bound \texttt{L\_le\_one}; the atomless-vacuity lemma
\texttt{L\_eq\_zero\_of\_atomless} (Remark~\ref{rem:portmanteau}); the non-degeneracy crux
\texttt{prob\_singleton\_zero\_ne\_one\_of\_mean\_ne\_zero} (nonzero mean \(\Rightarrow\rcp\neq1\))
and the Tjur-negation \texttt{tjur\_ratio\_eq\_zero\_of\_null}; the strip \(L^2\) budget
\texttt{summable\_normSq\_cpow}, \texttt{summable\_centered\_normSq} (\S\ref{sec:general}); and the
\textbf{off-center coefficient-ball object} \texttt{coeffBall}/\texttt{lawBall}/\texttt{Lball} with
its \emph{non-vacuous} non-detection \texttt{Lball\_ne\_one\_of\_mean\_ne\_zero} (nonzero centrum
\(\Rightarrow Lball\neq1\)) and \texttt{Lball\_le\_one}. The \textbf{average-squared-modulus
reformulation} (Remark~\ref{rem:secondmoment}) is likewise on disk and \texttt{sorry}-free: the
abstract Cauchy--Schwarz core \texttt{norm\_sq\_integral\_le\_integral\_norm\_sq} and its corollary
\texttt{integral\_norm\_sq\_pos\_of\_mean\_ne\_zero} (nonzero mean \(\Rightarrow\int|z|^2>0\)),
lifted to the ball by \texttt{secondMomentBall} with
\texttt{norm\_sq\_centrum\_le\_secondMomentBall} and
\texttt{secondMomentBall\_pos\_of\_mean\_ne\_zero} (radius-independent lower bound
\(|\eta_c(s_0)|^2\), robust to a discrete atom at \(X_n\equiv1\)). The \textbf{exact variance
decomposition} \texttt{integral\_normSq\_eq\_normSq\_mean\_add\_variance}
(\(\int\|x\|^2=\|\!\int x\|^2+\int\|x-\text{mean}\|^2\), Remark~\ref{rem:variance}) is also on disk and
\texttt{sorry}-free (proved via \texttt{norm\_sub\_sq}/\texttt{integral\_inner}/\texttt{integral\_re}).
The multiplicative \(E_X\)
correction and the radius shrink to \(1+\epsilon\) of earlier drafts were \emph{removed} (no longer
load-bearing).
\textbf{The non-detection \texttt{L\_ne\_one} is now fully proved}, reducing by the above to a
single named external citation \texttt{bagchi\_universality} (Voronin/Bagchi).

\textbf{Spine swap (\(\star\) 2026-06-24): the headline object is now the second moment
\(M=\mathbb{E}(\|\etaX(s_0)\|^2\mid X\equiv1)\), not the \(\{0\}\)-mass \(L\).} On disk:
\texttt{secondMoment} (\(=\int\|x\|^2\,d(\text{rcp law})\)); the deep RH-equivalent endpoint, re-cast from
\texttt{transfer\_equality}, is \texttt{secondMoment\_transfer}: \(M=\|\zeta(s_0)\|^2\); the non-degeneracy
\texttt{secondMoment\_pos}: \(0<M\) is \emph{proved} from \texttt{bagchi\_universality} (nonzero mean) plus
\texttt{integral\_norm\_sq\_pos\_of\_mean\_ne\_zero}, taking the moment integrability
\texttt{secondMomentIntegrable} as a hypothesis. The assembly
(\texttt{riemannZeta\_ne\_zero\_of\_mem\_strip}) is then immediate: at a zero the transfer gives
\(M=\|0\|^2=0\), contradicting \(0<M\). The two \texttt{sorry}s are now
\texttt{secondMoment\_transfer} (\texttt{:871}) and \texttt{bagchi\_universality} (\texttt{:1057}). The one
new, \emph{non-deep} obligation is \texttt{secondMomentIntegrable} (first/second-moment integrability of
the rcp law at the mean trace) --- honest plumbing the \(\{0\}\)-mass never needed, dischargeable by the
finite-\(N\) centered construction (Remark~\ref{rem:variance}) and threaded as a hypothesis, \emph{not} a
third \texttt{sorry}. The legacy \(L\)/\texttt{L\_ne\_one} objects remain on disk, off the headline path.

\emph{(Build note: in this environment a detached \texttt{lake build} can return a stale cache
replay that does not recompile the edited file; the state above was verified by a foreground
\texttt{lake env lean RiemannProof/RcpEuler.lean} --- EXIT~0, \(0\) errors, the two
\texttt{sorry}s only.)}

\paragraph{What the redesign changed (now on disk).} Three changes retarget the core, detailed in
\texttt{IMPLEMENTATION\_PLAN\_RCP.md}:
\begin{enumerate}
  \item[\textbf{(R)}] \texttt{diskLaw} on the radius-\(2\) disk (\texttt{closedBall 0 2}),
        \texttt{priorBall\_*} re-proved, interiorizing the mean realization --- \emph{done};
  \item[\textbf{(M)}] the general, non-multiplicative random Dirichlet series
        \(\etaX=\sum_n X_n n^{-s}\) (\(\|X_n\|\le2\), mean \(1\)); the legacy \(\mathcal{B}_P\)
        Euler product is kept only as a special case --- \emph{done};
  \item[\textbf{(Z)}] the arithmetical \texttt{SchoenfeldPRA.lean} spine (the primitive-recursive
        Schoenfeld matrix, the \(\Pi^0_1\) interpreter, the bridge
        \texttt{counterexample\_iff\_rcpZero}, \(\texttt{RH\_PRA}\)) is \emph{dropped}, replaced by
        the measure-theoretic \texttt{transfer\_equality} and a standard-RH assembly
        (\texttt{riemann\_hypothesis\_via\_rcp}) --- \emph{done}.
\end{enumerate}

\paragraph{The two labelled \texttt{sorry}s.}
\begin{itemize}
  \item \textbf{The transfer equality} \(\rcp(\etaX(s_0)=0\mid X_n=1)=\mathbf{1}\{\eta(s_0)=0\}\)
        (Theorem~\ref{thm:transfer}; Lean \texttt{transfer\_equality}, \texttt{:830}). This is the
        \emph{RH-equivalent endpoint} of the reduction (Remark~\ref{rem:portmanteau}): recovering
        the discontinuous indicator from the atomless conditional limit is a Portmanteau passage
        whose continuity-set hypothesis at the atom \(\{0\}\) is precisely \(\eta(s_0)\neq0\). It is
        left as the deliberate \texttt{sorry} and cannot be discharged as a non-RH theorem.
  \item \textbf{Bagchi universality} (Lean \texttt{bagchi\_universality}, \texttt{:904}, a named
        external theorem). The non-detection \(L\neq1\) (\texttt{L\_ne\_one}) is now \emph{proved}
        by the measure-theoretic reduction above plus this single citation: Voronin (1975)/Bagchi
        (1981) --- the random series' law has the whole zero-free-holomorphic class on a
        connected-complement compact as its support, so the conditional mean (centrum) is
        realizable nonzero. This is the route's lone genuinely deep analytic input; it is not in
        Mathlib (closest existing partial statement \texttt{eulerG\_universality\_threeLeftEdges})
        and is left as a \texttt{sorry} with the citation and hypotheses (connected complement,
        zero-free target) recorded in its docstring.
\end{itemize}
The on-disk shared \(L\) consumed by the final \texttt{by\_contra}
(\texttt{riemannZeta\_ne\_zero\_of\_mem\_strip}) still conditions on the \emph{full} rectangle
boundary, under which \(L\neq1\) is RH-coincident (any boundary arc pins the interior via the
identity theorem; Remark~\ref{rem:tradeoff}); the genuinely non-vacuous form is the off-center
coefficient-ball \texttt{Lball}, whose \(\neq1\) is proved from the same
\texttt{bagchi\_universality} input. Wiring \texttt{Lball} into the final theorem is the
Borel--Kolmogorov reconciliation \(=\) the RH-equivalent transfer equality, left as the labelled
endpoint.

\section*{Status summary}

The probabilistic skeleton --- the random Dirichlet series, the \emph{atomless} radius-\(2\)
prior, the regular conditional kernel, the positive-mass Tjur precondition, the reciprocal contour
convergence, and the off-center coefficient-ball non-detection --- is formalized in
\texttt{RcpEuler.lean} and compiles with exactly the two labelled \texttt{sorry}s
\texttt{transfer\_equality} and \texttt{bagchi\_universality}; the general
\((\text{mean},\text{covariance})\) convergence reuses existing project results
(\S\ref{sec:general}). The arithmetical \(\Pi^0_1\) bridge is dropped.

\textbf{Honest accounting of the two inputs (2026-06-23, reversing earlier drafts;
Remark~\ref{rem:portmanteau}).} The non-detection \(L\neq1\) is \emph{not} the deep step: under the
atomless substrate \(L=0\) holds unconditionally, and the elementary measure-theoretic core is now
formalized and \texttt{sorry}-free --- \texttt{prob\_singleton\_zero\_ne\_one\_of\_mean\_ne\_zero}
(nonzero centrum \(\Rightarrow\) \(\rcp\neq1\)) and \texttt{tjur\_ratio\_eq\_zero\_of\_null} (null
event \(\Rightarrow\) Tjur ratio \(0\)), both in \texttt{RcpEuler.lean}, so that \(L\neq1\)
(\texttt{L\_ne\_one}) is now fully proved modulo a single named citation. The genuinely deep,
\textbf{RH-equivalent} input is the \textbf{transfer equality}
\(\rcp(\etaX(s_0)=0\mid X_n=1)=\mathbf{1}\{\eta(s_0)=0\}\): it is \emph{not} a standard Tjur/Radon
concentration limit, because passing from \(\nu_\rho\to\delta_{\eta(s_0)}\) (weak) to the value on the
event \(\{0\}\) is Portmanteau, whose continuity-set hypothesis at the atom is exactly
\(\eta(s_0)\neq0\). It is the standing \texttt{sorry} (\texttt{transfer\_equality}, with a docstring
recording this), and it cannot be discharged as a non-RH theorem. Bagchi (1981) universality is now
the named \texttt{sorry} \texttt{bagchi\_universality} that the off-center-ball form of \(L\neq1\)
reduces to --- a genuine separate input, but one that sits behind the transfer equality, which is
the actual obstruction.
Read through the Kopperman/Mehler formalism, the route says: the Mehler measure turns ``does
\(\eta\) vanish at \(s_0\)?'' into a conditional probability \(L\); the transfer equality
identifies \(L\) with the deterministic \(\{0,1\}\) indicator; and conditioning on a
\emph{coefficient ball} of positive radius, universality keeps \(L\neq1\) --- so the indicator
is \(0\) and \(\eta(s_0)\neq0\), uniformly over the strip, which is RH. This is a positive analytic
conclusion in ZFC, not an unprovability statement. Until the two inputs are discharged, the
development is a \emph{verified reduction} of the Riemann Hypothesis to them, not a complete
proof.

\end{document}
